---
title: "PCRedux package - Case Study"
author: "The PCRedux package authors"
date: "`r Sys.Date()`"
output: 
  rmarkdown::pdf_document:
    number_sections: true
    toc: true
    toc_depth: 5
header-includes:
    - \usepackage[font={small}]{caption}
bibliography: "literature.bib"
---

```{r, include=FALSE, echo=FALSE, eval=FALSE}
options(tinytex.verbose = TRUE)
```


```{r, echo=FALSE}
knitr::opts_chunk$set(fig.lp = "", out.extra='', warnings = FALSE, message = FALSE)

amplification_curve_ROI <- "Phases of amplification curves as Region of Interest (ROI). For amplification curves, the fluorescence signal (RFU, relative fluorescence units) of the reporter dye is plotted against the cycle number. Positive amplification curves possess three ROIs: ground phase, exponential phase and plateau phase. These ROIs can be used to determine predictors such as the takedown point (`tdp`) or the standard deviation within the ground phase (`sd\\_bg`). The exponential range (red dots) is used to determine the Cq values and amplification efficiency (not shown). A linear regression model (red) can be used to calculate the slope in this region. B) PCRs without amplification reaction usually show a flat (non-sigmoides) signal. C) The exponential phase of PCR reactions can vary greatly depending on the DNA starting quantity and other factors. Amplification curves that appear in later cycles often have a lower slope in the exponential phase."

amplification_curve_ROI_short <- "Phases of amplification curves as Region of Interest (ROI)"

figure_quantification_points <- "Frequently used methods for the analysis of quantification points. A) The amplification curve is intersected by a gray horizontal line. This is the background signal (3$\\sigma$) determined from the \\textit{68-95-99.7 rule} from the fluorescence emission of cycles 1 to 10. The black horizontal line is the user-defined threshold (Ct value) in the exponential phase. Based on this, the cycle at which the amplification curve differs significantly from the background is calculated. B) The amplification curve can also be analyzed by fitting a multi-parametric model (black line, five parameters). The red line is the first derivative of the amplification curve with a maximum of 17.59 cycles. The first derivative maximum (`cpD1`) is used as a quantification point (Cq value) in some qPCR systems. The green line shows the second derivative of the amplification curve, with a maximum at 15.68 cycles a minimum at 19.5 cycles. The maximum of the second derivative (`cpD2`) is used as the Cq value in many systems. The blue line shows the amplification efficiency estimated from the trajectory of the exponential region. The `Eff` value of 1.795 means that the amplification efficiency is approximately 89\\%. `cpDdiff` is the difference between the first and second derivative maximum ($cpDdiff = cpD1 - cpD2$)."

figure_quantification_points_short <- "Frequently used methods for the analysis of quantification points"

figure_curve_classification <- "Variations in the classification of amplification curves. A prerequisite for the development of machine-learning models is the availability of manually classified amplification curves. Amplification curves (n = 8858) from the `htPCR` data set have been classified by one user eight times at different points over time (classes: ambiguous (a), positive (y) or negative (n)). During this process, the amplification curves were presented in random order. The example shows that different (subjective) class mappings may occur for the same data set. While only a few amplification curves were classified as negative in the first three classification cycles (A-C), their proportion increased almost tenfold in later classification cycles (D-H)."

figure_curve_classification_short <- "Variations of the classification of amplification curves"

htPCR_nap <- "Examples of negative, ambiguous and positive amplification curves. A) A negative (black), ambiguous (red) and positive (green) amplification curve were selected from the `htPCR` data set. The negative amplification curve is non-sigmoid and has a positive trend. The ambiguous amplification curve is similar to a sigmoidic amplification curve, but shows a positive slope in ground phase (cycle 1 $\\rightarrow$ 5). The positive amplification curve (green) is sigmoid. It starts with a flat baseline (cycle 5 $\\rightarrow$ 25). This is followed by the exponential phase (cycle 5 $\\rightarrow$ 25) and ends in a flat plateau phase (cycle 26 $\\rightarrow$ 35). B) Amplification curves of the `vermeulen1` data set were divided into groups with \\textit{negative}, \\textit{ambiguous} and \\textit{positive} classification. Negative amplification curves have a low signal level. Interesting is the spontaneous increase (probably due to a sensor calibration) in cycles 1 to 2 followed by a linear signal decrease. In principle, the ambiguous amplification curves have a sigmoid curve shape. However, the plateau phase is fairly broad. One of the ambiguous amplification curves begins to rise sharply at Cycle 45. The positive amplification curves have a characteristic sigmoid curve shape."

htPCR_nap_short <- "Examples of negative, ambiguous and positive amplification curves"

htPCR_nap_frequency <- "Frequency of amplification curve classes and conformity in the `htPCR` data set. The `htPCR` data set was classified by hand eight times. Due to the unusual amplification curve shape and input errors during classification, many amplification curves were classified differently. A) Frequency of negative (black), ambiguous (red) and positive (green) amplification curves in the `htPCR` data set. The combined number of ambiguous and negative amplification curves appears to be higher, than the number of positive amplification curves. B) The number of observations where all classification cycles made the same decision (conformity == TRUE) accounts for only 5\\% of the total number of observations. TRUE, all classes of the amplification curve matched. FALSE, at least one in eight observations had a different class."

htPCR_nap_short_frequency <- "Frequency of amplification curve classes and conformity in the `htPCR` data set."

qPCR2fdata <- "Shape-based clustering of amplification curves. A) The clustering of the amplification curves of the `testdat` data set (A) was based on the Hausdorff distance. B) The amplification curves were converted with the qPCR2fdata() function, and the Hausdorff distance of the curves was determined by cluster analysis. There were no errors in distinguishing between negative (n) and positive (y) amplification curves."

qPCR2fdata_short <- "Shape-based grouping of amplification curves"

HCU32 <- "Clustering and variation analysis of amplification curves. The amplification curves of the 32HCU were processed with the qPCR2fdata() function and then processed by cluster analysis (Hausdorff distance). A) Amplification curves were plotted from the raw data. B) Overall, signal-to-noise ratios of the amplification curves between all cavities were similar. C) The Cq values and amplification efficiency were calculated using the efficiency(pcrfit()) [\\texttt{qpcR}] function. The median Cq is shown as a vertical line. Cqs greater or less than 0. 1 of Cq $\\tilde{x}$ are marked with observation labels. D) The cluster analysis showed no specific pattern with respect to the amplification curve signals. It appears that the observations D1, E1, F1, F3, G3 and H1 differ most from the other amplification curves."

HCU32_short <- "Clustering and variation analysis of amplification curves"

winklR_principle <- "Concept of the winklR() function. Analysis of the amplification curves of the `RAS002` data set with the winklR() function. Two amplification curves (A: positive, B: negative) were used. The red point shows the origin (first negative derivative maximum) while the green and blue points show the minimum and maximum of the second negative derivative. The angle is calculated from these points. Positive curves have smaller angles than negative curves."

winklR_principle_short <- "Concept of the winklR() function"

winklR <- "Analysis of the amplification curves of the `RAS002` data set with the winklR() function. All amplification curves of the data set `RAS002` were analyzed. Negative amplification curves are shown in red and positive amplification curves in black. The winklR() function was used to analyze all amplification curves. B) The stripchart of the analysis of positive and negative amplification curves shows separation. C) The cdplot calculates the conditional densities of x based on the values of y weighted by the boundary distribution of y. The densities are derived cumulatively via the values of y. The probability that the decision is negative (n) when the angle equals 30 is approximately 100\\%."

winklR_short <- "Variation analysis of amplification curves with the winklR() function"

curve_fit_fail <- "Incorrect model adjustment for amplification curves. A positive (black), and a negative (red) amplification curve were randomly selected from the `RAS002` data set. The positive amplification curve has a baseline signal of about 2500 RFU and has a definite sigmoidal shape. The negative amplification curve has a baseline signal of approx. 4200 RFU, but only moderately positive slope (no sigmoidal shape). A logistic function with seven parameters (`l7`) has been fitted to both amplification curves. A Cq value of 25.95 was determined for the positive amplification curve. The negative amplification curve had a Cq value of 9.41. However, it can be seen that the latter model fitting is not appropriate for calculating a trustworthy Cq value. An automatic calculation without user control would give a false-positive result."

curve_fit_fail_short <- "Incorrect model adjustment for amplification curves"

plot_models <- "Frequencies of the fitted multiparametric models and Cq values. The amplification curves (n = 3302) of the `data\\_sample` data set were analyzed with the encu() function. The amplification curves have been stratified according to their classes (negative: grey, positive: green). A) The optimal multiparametric model was selected for each amplification curve based on the Akaike information criterion. lNA stands for `no model` and l4 \\ldots l7 for a model with four to seven parameters. B) All Cq values were calculated from optimal multiparametric models. Cqs of positive amplification curves accumulate in the range between 15 and 30 PCR cycles (50\\%). For the negative amplification curves, the Cqs are distributed over the entire span of the cycles. Note: The Cqs of the negative amplification curves are false-positive!"

plot_models_short <- "Frequencies of the fitted multiparametric models and Cq values"

figure_cpD2_range <- "Location of the predictors `cpD2\\_range`, `bg.start`, `bg.stop` within an amplification curve. The minimum (cpD2m) and maximum (cpD2) of the second derivative were calculated numerically using the diffQ2() function. This function also returns the maximum of the first derivative (cpD1). The `cpD2\\_range` is defined as $cpD2\\_range = |cpD2 - cpD2m|$. Large `cpD2\\_range` values indicate a low amplification efficiency or a negative amplification reaction. The predictor `bg.start` is an estimate for the end of the ground phase. `bg.start` is an approximation for the onset of the plateau phase."

figure_cpD2_range_short <- "Location of the of the predictors `cpD2\\_range`, `bg.start`, `bg.stop` within an amplification curve"

plot_dat_EffTop <- "Values of predictors calculated from negative and positive amplification curves. Amplification curve predictors from the `data\\_sample\\_subset` data set were used as they contain positive and negative amplification curves, and amplification curves that exhibit a \\textit{hook effect} or non-sigmoid shapes. A) `eff`, optimized PCR efficiency found within a sliding window. B) `sliwin`, PCR efficiency by the `window-of-linearity` method. C) `cpDdiff`, difference between the Cq values calculated from the first and the second derivative maximum. D) `loglin\\_slope`, slope from the cycle at the second derivative maximum to the second derivative minimum. E) `cpD2\\_range`, absolute value of the difference between the minimum and the maximum of the second derivative maximum. F) `top`, takeoff point. G) `f.top`, fluorescence intensity at takeoff point. H) `tdp`,  takedown point. I) `f.tdp`, fluorescence intensity at takedown point. J) `bg.stop`, estimated end of the ground phase. K) `amp.stop`, estimated end of the exponential phase. L) `convInfo\\_iteratons`, number of iterations until convergence."

plot_dat_EffTop_short <- "Analysis of location predictors"

loglin_slope <- "Concept of the `loglin\\_slope` predictor. The algorithm determines the fluorescence values of the raw data at the approximate positions of the maximum off the first derivative, the minimum of the second derivative and the maximum of the second derivative, which are in the exponential phase of the amplification curve. The data were taken from the `RAS002` data set. A linear model is created from these parameter sets and the slope is determined. A) Positive amplification curves have a clearly positive slope. B) Negative amplification curves usually have a low, sometimes negative slope."

loglin_slope_short <- "Concept of the `loglin\\_slope` predictor"

plot_sd_bg <- "Standard deviation in the ground phase of various qPCR devices. The `sd\\_bg` predictor was used to determine if the standard deviation between thermo-cyclers and between positive and negative amplification curves was different. The standard deviation was determined from the fluorescence values from the first cycle to the takeoff point. If the takeoff point could not be determined, the standard deviation from the first cycle to the eighth cycle was calculated. The Mann-Whitney test was used to compare the medians of the two populations (y, positive; n, negative). The differences were significant for A) LC\\_480 (Roche), B) CFX96 (Bio-Rad) and C) LC96 (Roche)."

plot_sd_bg_short <- "Standard deviation in the ground phase of various qPCR devices"

plot_bg_pt <- "Values of predictors calculated from negative and positive amplification curves. Amplification curve predictors from the `data\\_sample\\_subset` data set were used as they contain positive and negative amplification curves, as well as amplification curves that exhibit a \\textit{hook effect} or non-sigmoid shapes. A) `eff`, optimized PCR efficiency in a sliding window. B) `sliwin`, PCR efficiency according to the window-of-linearity method. C) `cpDdiff`, difference between the Cq values calculated from the first and the second derivative maximum. D) `loglin\\_slope`, slope from cycle at second derivative maximum to second derivative minimum. E) `cpD2\\_range`, absolute difference between the minimum and maximum of the second derivative. F) `top`, takeoff point. G) `f.top`, fluorescence intensity at takeoff point. H) `tdp`, takedown point. I) `f.tdp`, fluorescence intensity at the takedown point. J) `bg.stop`, estimated end of the ground phase. K) `amp.stop`, estimated end of the exponential phase. L) `convInfo\\_iteratons`, number of iterations until convergence when fitting a multiparametric model. The classes were compared using the Wilcoxon Rank Sum Test."

plot_bg_pt_short <- "Values of predictors calculated from negative and positive amplification curves"

plot_model_param <- "Values of predictors calculated from negative and positive amplification curves. Amplification curve predictors from the `data\\_sample\\_subset` data set were used as they contain positive and negative amplification curves, as well as amplification curves that exhibit a \\textit{hook effect} or non-sigmoid shapes. A) `c\\_model\\_param`, is the c model parameter of the seven parameter model. B) `d\\_model\\_param`, is the d model parameter of the seven parameter model. C) `e\\_model\\_param`, is the e model parameter of the seven parameter model. D) `f\\_model\\_param`, is the f model parameter of the seven parameter model. The classes were compared using the Wilcoxon Rank Sum Test."

plot_model_param_short <- "Values of predictors calculated from negative and positive amplification curves"

plot_Logistic_Regression <- "Machine classification by means of binomial logistic regression using the `loglin\\_slope` predictor. A) For the calculation of a binomial logistic regression model, the categorical response variable $Y$ (decision with classes: negative and positive) must be converted to a numerical value. With binomial logistic regression, the probability of a categorical response can be estimated using the $X$ predictor variable. In this example, the predictor variable `loglin\\_slope` is used. Grey measurement points (70\\% of the data set) were used for training. Red dots represent the values used for testing. The regression curve of the binomial logistic regression is shown in blue. The grey horizontal line at 0.5 marks the threshold of probability above which it is determined whether an amplification curve is negative or positive. B) The performance indicators were calculated using the performeR() function. Sensitivity, TPR; Specificity, SPC; Precision, PPV; Negative prediction value, NPV; Fall-out, FPR; False negative rate, FNR; False detection rate, FDR; Accuracy, ACC; F1 score, F1; Matthews correlation coefficient, MCC, Cohens kappa (binary classification), kappa ($\\kappa$)."

plot_Logistic_Regression_short <- "Machine classification by means of binomial logistic regression using the `loglin\\_slope` predictor"

statistical_methods_amptester <- "Analysis of amplification curves with the ``amptester()`` function. A \\& B) The threshold test (THt) is based on the Wilcoxon ranksum test and compares 20\\% of the fluorescence values of the ground phase with 15\\% of the plateau phase. In the example, a significant difference ($p = 0.000512$) was found for the positive amplification curve. However, this did not apply to the negative amplification curve ($p = 0.621$). C \\& D) A Q-Q diagram is used to graphically compare two probability distributions.  In this study the probability distribution of the amplification curve was compared with a theoretical normal distribution. The orange line is the theoretically normal quantil-quantile plot that passes through the probabilities of the first and third quartiles. The Shapiro-Wilk test (SHt) of normality checks whether the underlying measurement data of the amplification curve is significantly normal distributed. Since the p-value of $7.09 e^{-9}$ of the positive amplification curve is $\\alpha \\leq 5e^{-4}$, the null hypothesis is rejected. However, this does not apply to the negative amplification curve ($p = 0.895$). E \\& F) The linear regression test (LRt) calculates the coefficient of determination ($R^{2}$) using an ordinary least square regression where all measured values are integrated into the model in a cycle-dependent manner. Experience shows that the non-linear part of an amplification curve has a $R^{2}$ smaller than 0.8, which is also shown in the example."

statistical_methods_amptester_short <- "Analysis of amplification curves with the amptester() function"

figure_autocorrelation_tau <- "Effect of tau"

figure_autocorrelation_tau_short <- "Effect of tau"

autocorrelation <- "Autocorrelation analysis of the amplification curves of the `RAS002` data set. A) Display of all amplification curves of the data set `RAS002`. Negative amplification curves are shown in red and positive amplification curves in black. The autocorrelation\\_test() function was used to analyze all amplification curves. B) The density diagram of the autocorrelation of positive and negative amplification curves shows a bimodal distribution. C) The cdplot calculates the conditional densities of x based on the values of y weighted by the boundary distribution of y. The densities are derived cumulatively via the values of y. The probability that the decision is negative (n) when autocorrelation equals 0.85 is approximately 100\\%. D) Performance analysis using the performeR() function (see \\autoref{section_performeR} for details)."

autocorrelation_short <- "Autocorrelation analysis for amplification curves of the `RAS002` data set)"

earlyreg_slopes <- "Analysis of the ground phase with the earlyreg() function and the `C127EGHP` data set (n = 64 amplification curves). This data set consists of 32 samples, which were simultaneously monitored with the intercalator EvaGreen or hydrolysis probes. A) All amplification curves possess slightly different slopes and intercepts in the first cycles of the ground phase (ROI: Cycles 1 to 8). Both the slope and the intercept of each amplification curve were used for cluster analysis (k-means, Hartigan-Wong algorithm, number of centers \\textit{k = 2}). B) The amplification curves were assigned to five clusters, depending on their slope and their intersection (red, black). C) Finally, the clusters were associated to the detection chemistries (EvaGreen (EG) or hydrolysis probes (HP))."

earlyreg_slopes_short <- "Analysis of the ground phase with the earlyreg() function"

figure_head2tailratio <- "Ratio between the head and the tail of a quantitative PCR amplification curve. A) Plot of quantile normalized amplification curves from the `RAS002` data set. Data points used in the head and and tail are highlighted by circles. The intervals for the Robust Linear Regression are automatically selected using the 25\\% and 75\\% quantiles. Therefore, not all data points are used in the regression model. The straight line is the regression line from the robust linear model. The slopes of the positive and negative amplification curves differ. B) Boxplot for the comparison of the $head/tail$ ratio. Positive amplification curves have a lower ratio than negative curves. The difference between the classes is significant."

figure_head2tailratio_short <- "Ratio between the head and the tail of a quantitative PCR amplification curve"

plot_mblrr <- "Robust local regression to analyze amplification curves. The amplification curves were arbitrarily selected from the `RAS002` data set. In the qPCR setup, the target genes beta globin (B. globin) and HPRT1 were simultaneously measured in a PCR cavity using two specific hydrolysis probes (duplex qPCR). Both positive (A, C, E) and negative (B, D, F) amplification curves were used. The amplification curves are normalized to the 99\\% quantile. The differences in slopes and intercepts (blue and orange lines and dots). The mblrr() function is presumably useful for data sets which are accompanied by noise and artifacts."

plot_mblrr_short <- "Robust local regression to analyze amplification curves"

plot_FFTrees <- "Visualization of decisions in  Fast and Frugal Trees after data analysis of amplification curves via the mblrr() function. \\textbf{Top row} `Data`) Overview of the data set, stating the total number of observations (N = 192) and percentage of positive (22\\%) and negative (78\\%) amplification curves. \\textbf{Middle row} `FFT \\#1 (of 6)`) Decision Tree with the number of observations classified at each level of the tree. For the analysis, six predictors (nBG, intercept of head region; mBG, slope of head region; rBG, Pearson correlation of head region; nTP, intercept of tail region; mTP, slope of tail region; rBG, Pearson correlation of tail region) have been used for the analysis. After two tree levels (nBG, nTP), the decision tree is created, where all positive amplification curves (N = 40) are correctly classified. Two observations are classified as false-negative in the negative amplification curves. \\textbf{Lower row} `Performance`)  The FFTrees() [FFTrees] function determines several performance statistics. For the training data, there is a classification table on the left side showing the relationship between tree `decision` and the `truth`. The correct rejection (`Cor Rej`) and `Hit` are the right decisions. `Miss` and false alarm (`False Al`) are wrong decisions. The centre shows the cumulative tree performance in terms of mean of used cues (`mcu`), Percent of ignored cues (`pci`), sensitivity (`sens`), specificity (`spec`), accuracy (`acc`) and weighted Accuracy (`wacc`). The receiver operating characteristic (ROC) curve on the right-hand side compares the performance of all trees in the FFTrees object. The system also displays the performance of the fast frugal trees (`\\#`, green), CART (`C`, red), logistical regression (`L`, blue), random forest (`R`, violet) and the support vector machine (`S`, yellow)."

plot_FFTrees_short <- "Visualization of decisions in  Fast and Frugal Trees after data analysis of amplification curves via the mblrr() function"

plot_peaks_ratio <- "Working principle of the `peaks\\_ratio` predictor. The computation is based on a sequential linking of functions. The diffQ() function determines numerically the first derivative of an amplification curve. This derivative is passed to the mcaPeaks() [\\texttt{MBmca}] function. In the output all minima and all maxima are contained. The ranges are calculated from the minima and maxima. The Lagged Difference is determined from the ranges of the minima and maxima. Finally, the ratio of the differences (maximum/minimum) is calculated."

plot_peaks_ratio_short <- "Working principle of the `peaks\\_ratio` predictor"

plot_cp_area <- "Values of predictors calculated from negative and positive amplification curves. Amplification curves predictors from the `data\\_sample\\_subset` data set were used as they contain positive and negative amplification curves and amplification curves that exhibit a \\textit{hook effect} or non-sigmoid shapes. A) `polyarea`, is the area under the amplification curve determined by the Gauss polygon area formula. B) `peaks\\_ratio`, is the ratio of the local minima and the local maxima. C) `cp\\_e.agglo`, makes use of energy agglomerative clustering. Positive amplification curves have fewer change points than negative amplification curves. These two change point analyses generally separate positive and negative amplification curves. D) `cp\\_bcp`, analyses change points by a Bayesian approach. Positive amplification curves appear to contain more change points than negative amplification curves. Nevertheless, there is an overlap between the positive and negative amplification curves in both methods. This can lead to false-positive or false-negative classifications. E) `amptester\\_polygon` is the cycle normalized order of a polygon.  F) `amptester\\_slope.ratio` is the slope (linear model) of the raw fluorescence values at the approximate first derivate maximum, second derivative minimum and second derivative maximum."

plot_cp_area_short <- "Analysis of predictors that decribe the area and changepoints of an amplification curve"

plot_cpa <- "Bayesian and energy agglomerative change point analysis on negative and positive amplification curves. An analysis of a negative and a positive amplification curve from the `RAS002` data set was performed using the pcrfit\\_single() function. In this process, the amplification curves were analysed for change points using Bayesian change point analysis and energy agglomerative clustering. A) The negative amplification curve has a base signal of approximately 2450 RFU and only a small signal increase to 2650 RFU. There is a clear indication of the signal variation (noise). B) The first negative derivative amplifies the noise so that some peaks are visible. C) The change point analysis shows changes in energy agglomerative clustering at several positions (green vertical line). The Bayesian change point analysis rarely exceeds a probability of 0.6 (grey vert line). D) The positive amplification curve has a lower base signal ($\\sim 2450$ RFU) and increases up to the 40th cycle ($\\sim 3400$ RFU). A sigmoid shape of the curve is visible. E) The first negative derivation of the positive amplification curve shows a distinctive peak with a minimum at cycle 25. F) The change point analysis in energy agglomerative clustering shows changes (green vertical line) only at two positions. The Bayesian change point analysis shows a probability higher than 0.6 (grey horizontal line) at several positions."

plot_cpa_short <- "Bayesian and energy agglomerative change point analysis on negative and positive amplification curves"

plot_random_forest <- " The predictors `amptester\\_lrt` (lrt), `amptester\\_rgt` (rgt), `amptester\\_tht` (tht), `amptester\\_slt` (slt), `amptester\\_polygon`(polygon) and `amptester\\_slope.ratio` (slope.ratio) were used for classification using random forest. A) This plot shows the error depending on the number of trees. The error decreases as more and more trees are added and averaged. B). Mean Decrease Accuracy shown how much the model accuracy decreases if a variable is dropped. C) Mean Decrease Gini shows the importance of a variable based on the Gini impurity index used for the calculation of splits in trees."

plot_random_forest_short <- "Random Forest"


#----------Tables-------------------------------------
```

\newpage

\begin{figure}[ht]
\centering
    \scalebox{0.6}{
    \includegraphics[clip=true,trim=1cm 1cm 1cm 1cm]{Logo.pdf}
}
\end{figure}


# Case study for the application of the PCRedux package

In this case study we used 400 amplification curves from the *kbqPCR* dataset 
from and assign them to the object *curves*. A comprehensive PDF version 
(including domain knowledge about qPCRs and machine learning) of this document 
is available **[online 
supplement](https://github.com/devSJR/PCRedux/raw/master/docs/articles/PCRedux.pdf)**. This online document contains additional code (incl. installation of the 
software, data import, preprocessing)) that helps to understand how this 
introductionary case study was created. Before we start we need to load 
additional libraries (some from CRAN, some from github.com).

```{R}
library(qpcR)
library(PCRedux)

library(dplyr)
library(forcats)
library(reshape2)

library(mlr)

library(ggplot2)

if("devtools" %in% rownames(installed.packages()) == FALSE) {
  library(devtools)
}

if("gbm" %in% rownames(installed.packages()) == FALSE) {
  install.packages("gbm")
}

if("patchwork" %in% rownames(installed.packages())) {
  library(patchwork)
} else {
  devtools::install_github("thomasp85/patchwork")
  "patchwork" %in% rownames(installed.packages())
  library(patchwork)
}
```

For the example qPCR data from a real experiment with a CFX96 (Bio-Rad) were used. It is an allelic specific PCR with TaqMan probes (FAM, HEX, ROX, Cy5, Cy5-5) (experimental details are described in @romaniuk_rapid_2019. Original pcrd and rdml files are here:
[https://github.com/kablag/AScall/tree/master/examples](https://github.com/kablag/AScall/tree/master/examples)

```{R}
# Determine the Number of amplification curves
curves <- ncol(kbqPCR)

paste("Number of amplification curves:", curves - 1)
```

The decision of the classes (negative, ambiguous, positive) were taken from the `decision_res_kbqPCR.rda` of the *PCRedux* package.

```{r}
dec <- unlist(lapply(1L:(curves -1), function(i) {
  decision_modus(decision_res_kbqPCR[i, 2:8])
}))

class <- c(y = sum(dec == "y"), a = sum(dec == "a"), n = sum(dec == "n"))

# Plot the classes
barplot(table(dec), col = c("green", "red", "grey"), 
        main = "Frequency of Classes assed by a Human")
legend("topright", c("y: positive", "n: negtive", "a: ambigous"))
text(c(1:3), rep(150,3), class[c(1,3,2)])
```

## Visualizng qPCR curves separated as positive, negative and ambigous Amplification Curves

As a definition for a positive amplification curve, it is assumed that it has a sigmoid shape and a good signal-to-noise ratio. An ambiguous amplification curve has a signal that is above the noise level but has no sigmoid form. In this example, the ambiguous amplification curves are almost linear and do not show a plateau. Negative amplification curves do not achieve a high signal and do not show a sigmoid curve.


```{r}
p1_pos <- data.frame(kbqPCR[, c(1L, which(dec == "y") + 1)]) %>% 
  melt(id.vars = "cyc") %>%
  ggplot(aes(x = cyc, y = value, color = variable)) +
  geom_line() +
  theme_bw() +
  xlab("Cycle") +
  ylab("Raw RFU") +
  ggtitle(paste0("A) positive, ", "n = ", class[1])) + #qPCR curves
  theme(legend.position = "none")

# Positive Curves
p1_pos

p1_neg <- data.frame(kbqPCR[, c(1L, which(dec == "n") + 1)]) %>% 
  melt(id.vars = "cyc") %>%
  ggplot(aes(x = cyc, y = value, color = variable)) +
  geom_line() +
  theme_bw() +
  xlab("Cycle") +
  ylab("Raw RFU") +
  ggtitle(paste0("negative, ", "n = ", class[3])) + #qPCR curves
  theme(legend.position = "none")

# Negative Curves
p1_neg

p1_amb <- data.frame(kbqPCR[, c(1L, which(dec == "a") + 1)]) %>% 
  melt(id.vars = "cyc") %>%
  ggplot(aes(x = cyc, y = value, color = variable)) +
  geom_line() +
  theme_bw() +
  xlab("Cycle") +
  ylab("Raw RFU") +
  ggtitle(paste0("ambiguous, ", "n = ", class[2])) + #qPCR curves
  theme(legend.position = "none")

# Ambiguous Curves
p1_amb

```


## Calculating some parameters with the encu() function

The *encu()* function was used to calculate some parameters that are later used for the machine learning. Please be patient, this step will take some time.

```{r}
res <- encu(kbqPCR[, 1L:curves])


head(res)
```

## Merging decisions and features into one dataset

Since the parameters from the calculations with the *encu()* function and the decisions are known by now, we can merge them into a single dataframe.

```{r}
dat <- cbind(res, decision = factor(c("ambiguous", "negative", "positive")[dec], 
                                    levels = c("positive", "ambiguous", "negative"))) %>%
  select(cpD2, amptester_polygon, cpDdiff, decision) %>%
  filter(!is.na(dec))

head(dat)
```

## Visualizing the parameter calculations

Next we visualize the results of the parameter calculations.

```{r}
p2 <- ggplot(data = dat %>%
               mutate(id = rownames(dat)) %>%
               melt(id.vars = c("id", "decision")), 
             aes(x = decision, y = value)) +
  geom_boxplot() +
  theme_bw() +
  facet_wrap(~ variable, scales = "free_y") +
  scale_y_continuous("Value [A.U.]") +
  scale_x_discrete("Assessment") +
  ggtitle("B)") # Separation of types of curves by encu() parameters

p2
```

## Modeling with different classifiers

Building the models follows the typical steps involving the definition of the splits, classifiers and their tasks. Here we use the *mlr* package.
We want to classify only! As classifiers we use:

- Random Forest (classif.ranger),
- Support Vector Machines (classif.ksvm),
- linear discriminant analysis (classif.lda),
- Generalized Boosted Regression Models (classif.gbm),
- Multinomial Regression (classif.multinom) and
- Generalized Linear Regression with Lasso or Elasticnet Regularization (classif.glmnet).

```{r}
tsk <- makeClassifTask("pcr_classif", data = dat, target = "decision")

mdls <- list()
mdls[[1]] <- makeLearner("classif.ranger", predict.type = "prob")
mdls[[2]] <- makeLearner("classif.ksvm", predict.type = "prob")
mdls[[3]] <- makeLearner("classif.lda", predict.type = "prob")
mdls[[4]] <- makeLearner("classif.gbm", predict.type = "prob")
mdls[[5]] <- makeLearner("classif.multinom", predict.type = "prob")
mdls[[6]] <- makeLearner("classif.glmnet", predict.type = "prob")

set.seed(4732)
results <- do.call(rbind, lapply(mdls, function(mdl) {
  res <- resample(mdl, tsk, cv10, measures = list(mmce, multiclass.au1u))
  cbind(model = res[["learner.id"]], res[["measures.test"]])
}))
results[["model"]] <- fct_recode(results[["model"]],
                                 `ranger::ranger` = "classif.ranger", 
                                 `kernlab::ksvm` = "classif.ksvm",
                                 `MASS::lda` = "classif.lda",
                                 `gbm::gbm` = "classif.gbm",
                                 `nnet::multinom` = "classif.multinom",
                                 `glmnet::glmnet` = "classif.glmnet")
```

## Visualizing the model results

```{r}
p3 <- ggplot(data = results, aes(x = model, y = multiclass.au1u)) +
  geom_point(position = position_jitter(width = 0.2, seed = 4)) + 
  geom_errorbar(data = results %>% 
                  group_by(model) %>% 
                  summarise(auc = median(multiclass.au1u)), 
                aes(x = model, ymin = auc, ymax = auc), 
                inherit.aes = FALSE, color = "#FC5E61",
                width = 0.5) +
  theme_bw() + 
  xlab("Model") +
  ylab("Mean AUC (one vs all)") +
  ggtitle("C)") # Results of crossvalidating models trained on encu() parameters

p3
```

## Final plot

Finally we plot all findings in a summary graphic.

```{r}
cairo_ps("figure1.eps", width = 11, height = 3.1)
p1_pos + {
    p1_neg +
      p1_amb +
      plot_layout(ncol = 1)
  } + p2 + plot_layout(ncol = 3, widths = c(1, 1, 4))
dev.off()
```

# References

